---
title: "Class11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB Database

The [PDB](http://www.rcsb.org/) is the main respository for biomolectular structure data.

Here we examine the contents of the PDB:

> Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. From the website what proportion of structures are protein?

```{r}
db <- read.csv("Data Export Summary.csv", row.names=1)
head(db)
```

How many are X-Ray
```{r}
(db$Total/sum(db$Total))*100
```

```{r}
(sum(db$Proteins)/sum(db$Total)) * 100
```

We could also try the datapasta pacakge and copy from website and "Addins" > "Paste as data.frame"
```{r}
library(datapasta)

tmp <- data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(126880, 11062, 2277, 256, 129, 140604),
         Nucleic.Acids = c(2012, 1279, 31, 4, 5, 3331),
    ProteinComplex = c(6547, 259, 800, 6, 2, 7614),
                 Other = c(8, 8, 0, 13, 1, 30),
                 Total = c(135447, 12608, 3108, 279, 137, 151579)
)

```

> Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB? 

There are 1157 as of "2019-05-07"
See: http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=1DB596F9


```{r}
library(bio3d)

pdb <- read.pdb("1hsg.pdb")
pdb

```
```{r}
pdb$atom
```


Atom selection is done via the funtion **atom.select()**
```{r}
inds <- atom.select(pdb, "protein")
```

```{r}
pdb$atom[inds$atom, ]
```

#writes into protein only pdb file
```{r}
prot.pdb <- atom.select(pdb, "protein",value = TRUE)
write.pdb(prot.pdb, file="1hsg_protein.pdb")
```

```{r}
lig.pdb <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(lig.pdb, file="1hsg_ligand.pdb")
```

```{r}
aa <- get.seq("1ake_A")
b<- blast.pdb(aa)
```

```{r}
hits <- plot(b)
```

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
pdbs <- pdbaln(files)
ids <- basename.pdb(pdbs$id)
plot(pdbs, labels=ids)
```

```{r}
cons <- conserv(pdbs, method="entropy22")
sse <- pdbs2sse(pdbs, ind=1, rm.gaps=FALSE)
plotb3(cons, sse=sse, ylab="Sequence entropy")
```

```{r}
anno <- pdb.annotate(ids)
print(unique(anno$source))
```

```{r}
core <- core.find(pdbs)
pdbs$xyz = pdbfit(pdbs, core)
pc.xray <- pca(pdbs)
rd <- rmsd(pdbs)
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```


















