---
title: "Class9: Unsupervised Learning Analysis of Human Breast Cancer Cells"
author: "Jimin Park"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##PART 1 Exploratory Data Analysis
```{r}
fna.data <- "data/WisconsinCancer.csv"
wisc.df <- read.csv(fna.data)
#head(wisc.df)
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)
row.names(wisc.data) <- wisc.df$id
#Store the diagnosis data as a vector of 1 and 0 values
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```

> Q1. How many observations are in this dataset?
```{r}
length(wisc.data)
```

> Q2. How many variables/features in the data are suffixed with _mean?
```{r}
length(grep("_mean", colnames(wisc.data)))
```

> Q3. How many of the observations have a malignant diagnosis?
```{r}
sum(diagnosis)
```

##PART 2 Principal Component Analysis
```{r}
round(colMeans(wisc.data), 1)
```

```{r}
round(apply(wisc.data,2,sd), 1)
```

```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```

> Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)? 
44.27%

> Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data? 
PC3

> Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data? 
PC7

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1, xlab = "PC1", ylab = "PC2")
```

> Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots? 
The separation between malignant and benign is more obvious in the first plot since PC2 captures more variance than PC3.

```{r}
# Repeat for components 1 and 3
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC3")
```

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```
```{r}
# Variance explained by each principal component: pve
pve <- (pr.var / sum(pr.var))*100
head(pve)
```

#Make a scree plot
```{r}
# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```
```{r}
# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")
```
```{r}
par(mfcol=c(1,2))
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")

plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")
```
```{r}
## ggplot based graph
library(factoextra)
```

```{r}
fviz_eig(wisc.pr, addlabels = TRUE)
```

> Q9. For the first principal component, what is the component of the loading vector (i.e.  wisc.pr$rotation[,1]) for the feature concave.points_mean?

> Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?

##PART 3 Hierarchical Clustering
```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "complete")
```

> Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

```{r}
plot(wisc.hclust)
abline(h = 18, col="red", lty=2)
```

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
table(wisc.hclust.clusters, diagnosis)

```
> Q12. Can you find a better cluster vs diagnoses match with by cutting into a different number of clusters between 2 and 10?

##PART 5 Combining Methods: Clustering on PCA results
```{r}
dist.wisc.pr <- dist(wisc.pr$x[,1:7])
wisc.pr.hclust <- hclust(dist.wisc.pr, method="ward.D2")
plot(wisc.pr.hclust)
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis+1)
```

## PART 7 Prediction
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16)
```

> Q17. Which of these new patients should we prioritize for follow up based on your results?
We should prioritize the patient in the malignant cluster.c













