---
title: 'Class 7: R functions and packages'
author: "Jimin Park"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions revisited

We will source a file from online with our functions from previous class

```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale() function

```{r}
rescale(1:10)
```


Try the rescale2() funciton that catches string inputs
```{r}
rescale2(c(1:10, "string"))
```

# Find missing NA values in two vectors

Start with a simple example of the larger problem I am trying to solve

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```
```{r}
is.na(x)
is.na(y)
```

Try putting these together with an AND

```{r}
is.na(x) & is.na(y)
```

Take the sum() to find out how many TRUE values we have in both and thus how many NA's we have in both x and y
```{r}
sum(is.na(x) & is.na(y))
```

Now I can make this into our first function:
```{r}
both_na <- function(x, y){
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x, c(NA, 3, NA, 2, NA))
```

Test, test, test
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na(x, y2)
```
```{r}
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```

Now lets try the both_na2() function on our different length input vectors

```{r}
both_na2(x, y2)
```

```{r}
which(c(F, F, T, F, T))

which(is.na(c(1, 2, NA, 4)))
```

```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(1, NA, 2, 4, NA)

both_na3(x, y)
```

Make things simple
```{r}
x <- df1$IDs
y <- df2$IDs
```

```{r}
intersect(x, y)
```

```{r}
which(x %in% y)
```

```{r}
x[x %in% y]
y[y %in% x]
```

Use the RStudio shortcut 'CODE > EXTRACT FUCTION' to turn our snippet into a working function
```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y],
        y[y %in% x])
}
```

```{r}
gene_intersect(df1$IDs, df2$IDs)
```

```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1, df2)
```
```{r}
merge(df1, df2, by = )
```

Function "Grade": to calculate mean score drop the lowest single score

```{r}
student1 <- c(100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, 90, 90, 90, 90, 90, 97, 80)
```

```{r}
grade <- function(student){
  (sum(student) - min(student)) / (length(student) - 1)
}
```

```{r}
grade(student1)
grade(student2)
```























